#Modules
```{r}
library(ggplot2)
library(tsibble)
library(tseries)
library(MASS)
library(forecast)
library(TSA)
library(TTR)
library(aTSA)
library(graphics)
```


#Input data
```{r}
df=read.csv("https://raw.githubusercontent.com/Rheyhan/MPDW-Rhey/master/Tugas/Harga%20Saham%20Lockheed%20Martin%20-%20Sheet1.csv")
data.ts=ts(df$Close)

#Data partisi 1
df1=df[1:round(length(df$Date)*0.5),]
data1.ts=ts(df1$Close)
  
#Data partisi 2
temp=round(length(df$Date)*0.5)+1

df2=df[temp:length(df$Date),]
data2.ts=ts(df2$Close)
```
#Data Penuh
##Train test split
```{r}
#Train test split
train=df[1:round(length(df$Date)*0.8),]
train.ts=ts(train$Close)
temp=round(length(df$Date)*0.8)+1
test=df[temp:length(df$Date),]
test.ts=ts(test$Close)
```
##Exploration
```{r}
plot.ts(data.ts, lty=1, main="Entire data")
plot.ts(train.ts, lty=1, main="Train")
plot.ts(test.ts, lty=1, main="Test")
```
##Check for Stationary
```{r}
#Check stationer
plot(train.ts)    

  #plot acf
acf(train.ts)

  #plot pacf
pacf(train.ts)

  #plot eacf
eacf(train.ts)

  #Dickey-Fuller Test
tseries::adf.test(train.ts)     #Not stationary
```
##Handling not stationary
###Box cox
```{r}
index <- seq(1:length(train.ts))
bc = boxcox(train.ts~index, lambda = seq(1,10,by=1))
#Nilai Rounded Lambda
lambda <- bc$x[which.max(bc$y)]
lambda
#SK
bc$x[bc$y > max(bc$y) - 1/2 * qchisq(.95,1)]
```

###diff
```{r}
train.diff=diff(train.ts,differences = 1)
plot.ts(train.diff, lty=1, main="Plot Difference")
```
###Recheck
```{r}
  #plot acf
acf(train.diff)

  #Plot eacf
eacf(train.diff)

  #Dickey-Fuller Test
tseries::adf.test(train.diff)     #stationary
```
##Get model
```{r}
#Get arima model
model = auto.arima(train.diff, stationary = T, max.p =  10, max.Q =  10, max.d =  10, max.P =  10, max.D = 10,)    #Auto arima
model

model1 = Arima(train.diff, order=c(0,0,3),method="ML")
summary(model1) #AIC= 5098.23   
lmtest::coeftest(model1)
```
###Resisual Check
```{r}
sisaan.da = model1$residuals 
par(mfrow=c(2,2)) 
qqnorm(sisaan.da) 
qqline(sisaan.da, col = "blue", lwd = 2) 
plot(c(1:length(sisaan.da)),sisaan.da) 
acf(sisaan.da) 
pacf(sisaan.da) 
par(mfrow = c(1,1))
```
###Formal Testing
```{r}
#1) Sisaan Menyebar Normal 
ks.test(sisaan.da,"pnorm")    #Tak tolak H0

#2) Sisaan saling bebas/tidak ada autokorelasi 
Box.test(sisaan.da, type = "Ljung")   #Tolak H0

#3) Sisaan homogen 
Box.test((sisaan.da)^2, type = "Ljung") #Tak Tolak H0

#4) Nilai tengah sisaan sama dengan nol 
t.test(sisaan.da, mu = 0, conf.level = 0.95)    #Tolak H0
```
##Forecasting
```{r}
h=length(test.ts)
ramalan.da <- forecast::forecast(model1, h = h) 
data.ramalan.da <- ramalan.da$mean
plot(ramalan.da)
```
##
```{r}
pt_1 <- train.ts[length(train.ts)] #nilai akhir data latih
hasil <- diffinv(data.ramalan.da, differences = 1) + pt_1
ts.plot(train.ts,hasil)
```
```{r}
test.ts

perbandingan.da<-matrix(c(test.ts, hasil[-1]),
                     nrow = h, ncol = 2)
colnames(perbandingan.da)<-c("Aktual","Forecast")
perbandingan.da
accuracy(ts(hasil[-1]), head(test.ts, n=h))
```
#Data Partisi 1
##Train test split
```{r}
#Train test split
train=df1[1:round(length(df1$Date)*0.8),]
train.ts=ts(train$Close)
temp=round(length(df1$Date)*0.8)+1
test=df1[temp:length(df1$Date),]
test.ts=ts(test$Close)
```
##Exploration
```{r}
plot.ts(data1.ts, lty=1, main="Entire data")
plot.ts(train.ts, lty=1, main="Train")
plot.ts(test.ts, lty=1, main="Test")
```
##Check for Stationary
```{r}
#Check stationer
plot(train.ts)    

  #plot acf
acf(train.ts)

  #plot pacf
pacf(train.ts)

  #plot eacf
eacf(train.ts)

  #Dickey-Fuller Test
tseries::adf.test(train.ts)     #Not stationary
```
##Handling not stationary
###Box cox
```{r}
index <- seq(1:length(train.ts))
bc = boxcox(train.ts~index, lambda = seq(1,10,by=1))
#Nilai Rounded Lambda
lambda <- bc$x[which.max(bc$y)]
lambda
#SK
bc$x[bc$y > max(bc$y) - 1/2 * qchisq(.95,1)]
```

###diff
```{r}
train.diff=diff(train.ts,differences = 1)
plot.ts(train.diff, lty=1, main="Plot Difference")
```
###Recheck
```{r}
  #plot acf
acf(train.diff)

  #Plot eacf
eacf(train.diff)

  #Dickey-Fuller Test
tseries::adf.test(train.diff)     #stationary
```
##Get model
```{r}
#Get arima model
model = auto.arima(train.diff, stationary = T, max.p =  10, max.Q =  10, max.d =  10, max.P =  10, max.D = 10,)    #Auto arima
model

model1 = Arima(train.diff, order=c(0,0,0),method="ML")
summary(model1) #AIC= 2195.21   
lmtest::coeftest(model1)
```
###Resisual Check
```{r}
sisaan.da = model1$residuals 
par(mfrow=c(2,2)) 
qqnorm(sisaan.da) 
qqline(sisaan.da, col = "blue", lwd = 2) 
plot(c(1:length(sisaan.da)),sisaan.da) 
acf(sisaan.da) 
pacf(sisaan.da) 
par(mfrow = c(1,1))
```
###Formal Testing
```{r}
#1) Sisaan Menyebar Normal 
ks.test(sisaan.da,"pnorm")    #Tak tolak H0

#2) Sisaan saling bebas/tidak ada autokorelasi 
Box.test(sisaan.da, type = "Ljung")   #Tolak H0

#3) Sisaan homogen 
Box.test((sisaan.da)^2, type = "Ljung") #Tak Tolak H0

#4) Nilai tengah sisaan sama dengan nol 
t.test(sisaan.da, mu = 0, conf.level = 0.95)    #Tolak H0
```
##Forecasting
```{r}
h=length(test.ts)
ramalan.da <- forecast::forecast(model1, h = h) 
data.ramalan.da <- ramalan.da$mean
plot(ramalan.da)
```
##
```{r}
pt_1 <- train.ts[length(train.ts)] #nilai akhir data latih
hasil <- diffinv(data.ramalan.da, differences = 1) + pt_1
ts.plot(train.ts,hasil)
```
```{r}
test.ts

perbandingan.da<-matrix(c(test.ts, hasil[-1]),
                     nrow = h, ncol = 2)
colnames(perbandingan.da)<-c("Aktual","Forecast")
perbandingan.da
accuracy(ts(hasil[-1]), head(test.ts, n=h))
```
#Data Partisi 2
##Train test split
```{r}
#Train test split
train=df2[1:round(length(df2$Date)*0.8),]
train.ts=ts(train$Close)
temp=round(length(df2$Date)*0.8)+1
test=df2[temp:length(df2$Date),]
test.ts=ts(test$Close)
```
##Exploration
```{r}
plot.ts(data2.ts, lty=1, main="Entire data")
plot.ts(train.ts, lty=1, main="Train")
plot.ts(test.ts, lty=1, main="Test")
```
##Check for Stationary
```{r}
#Check stationer
plot(train.ts)    

  #plot acf
acf(train.ts)

  #plot pacf
pacf(train.ts)

  #plot eacf
eacf(train.ts)

  #Dickey-Fuller Test
tseries::adf.test(train.ts)     #stationary
```
##Get model
```{r}
#Get arima model
train.diff=diff(train.ts,differences = 1)

model = auto.arima(train.diff, stationary = T, max.p =  10, max.Q =  10, max.d =  10, max.P =  10, max.D = 10,)    #Auto arima
model
model1=Arima(train.diff, c(2,0,2))
summary(model1) #AIC= 3340.83         
lmtest::coeftest(model1)
```
###Resisual Check
```{r}
sisaan.da = model1$residuals 
par(mfrow=c(2,2)) 
qqnorm(sisaan.da) 
qqline(sisaan.da, col = "blue", lwd = 2) 
plot(c(1:length(sisaan.da)),sisaan.da) 
acf(sisaan.da) 
pacf(sisaan.da) 
par(mfrow = c(1,1))
```
###Formal Testing
```{r}
#1) Sisaan Menyebar Normal 
ks.test(sisaan.da,"pnorm")    #Tak tolak H0

#2) Sisaan saling bebas/tidak ada autokorelasi 
Box.test(sisaan.da, type = "Ljung")   #Tolak H0

#3) Sisaan homogen 
Box.test((sisaan.da)^2, type = "Ljung") #Tak Tolak H0

#4) Nilai tengah sisaan sama dengan nol 
t.test(sisaan.da, mu = 0, conf.level = 0.95)    #Tolak H0
```
##Forecasting
```{r}
h=length(test.ts)
ramalan.da <- forecast::forecast(model1, h = h) 
data.ramalan.da <- ramalan.da$mean
plot(ramalan.da)
```
##
```{r}
pt_1 <- train.ts[length(train.ts)] #nilai akhir data latih
hasil <- diffinv(data.ramalan.da, differences = 1) + pt_1
ts.plot(train.ts,hasil)
ts.plot(train.ts,hasil)
```
```{r}
test.ts

perbandingan.da<-matrix(c(test.ts, hasil[-1]),
                     nrow = h, ncol = 2)
colnames(perbandingan.da)<-c("Aktual","Forecast")
perbandingan.da
accuracy(ts(hasil[-1]), head(test.ts, n=h))
```






